Lab 2 Final Report: COVID-19 Case Rate vs Population Demographics and Mask Policy

### 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Update Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
update model 3

```{r variables, warning=FALSE, results='hide'}
df <- read.csv("covid-19.csv", header = TRUE)
df<-df%>%
  rename(case_rate_100k = 'Case.Rate.per.100000',
         death_rate_100k = 'Death.Rate.per.100000',
         population_density = 'Population.density.per.square.miles',
         white_pct = 'White...of.Cases',
         black_pct = 'Black...of.Cases',
         hispanic_pct = 'Hispanic...of.Cases',
         other_pct = 'Other...of.Cases',
         state_emergency = 'State.of.emergency',
         business_closed = 'Closed.other.non.essential.businesses',
         business_reopen = 'Began.to.reopen.businesses.statewide',
         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
         mask_legal='No.legal.enforcement.of.face.mask.mandate',
         black_population_pct ="Black...of.Total.Population",
         white_population_pct = "White...of.Total.Population",
         poverty_pct = "Percent.living.under.the.federal.poverty.line..2018.",
         unemployed_pct = "Percent.Unemployed..2018.",
         senior_pct = "X65.",
         population_2018 = "Population.2018", ## extra variable in test
         non_elderly_pre_existing = "Nonelderly.Adults.Who.Have.A.Pre.Existing.Condition",
         homeless = "Number.Homeless..2019.") %>%
  select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, 
         black_population_pct, white_population_pct, poverty_pct, unemployed_pct, senior_pct,
         population_2018, non_elderly_pre_existing, homeless)
# Assign correct data types
cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
df[cols.num] <- sapply(df[cols.num],as.numeric)
df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
head(df)
```

@@

#### Model 3 Regression

Next, the regression model for model 3 is based on the existing work from model 2. Two additional variables discussed in Model 3 will be introduced as maximalist approach. The first variable is the "mask_public_bool" (whether the state enforced public mask mandate or not) that represents the policy effects. The second variable introduced is "population_density" (the value is logged as discussed earlier) which is considered an important factor considering nature of spread in disease.

@@

```{r}
model3 <- lm(case_rate_100k ~ senior_per_100k 
              + poverty_per_100k 
              + unemployed_per_100k 
             + mask_public_bool
             + log(population_density), data = df)
coeftest(model3)

@@

CLM assumptions check for Model 3:

@@

#### Normality of errors
From the Normal Q-Q plot below, the points generally stay close to the normal line, with the tails exhibit certain degree of skew. 
```{r}
plot(model3, which=2)
```

#### Homoskedasticity
```{r}

plot(model3, which=3)
```
```{r}
#Cook distance
plot(model3, which=5)
```
```{r}
#Collinearity
vif(model3)
```

>>>>

```{r variables, warning=FALSE, results='hide'}
df <- read.csv("covid-19.csv", header = TRUE)
df<-df%>%
  rename(case_rate_100k = 'Case.Rate.per.100000',
         death_rate_100k = 'Death.Rate.per.100000',
         population_density = 'Population.density.per.square.miles',
         white_pct = 'White...of.Cases',
         black_pct = 'Black...of.Cases',
         hispanic_pct = 'Hispanic...of.Cases',
         other_pct = 'Other...of.Cases',
         state_emergency = 'State.of.emergency',
         business_closed = 'Closed.other.non.essential.businesses',
         business_reopen = 'Began.to.reopen.businesses.statewide',
         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
         mask_legal='No.legal.enforcement.of.face.mask.mandate',
         black_population_pct ="Black...of.Total.Population",
         white_population_pct = "White...of.Total.Population",
         poverty_pct = "Percent.living.under.the.federal.poverty.line..2018.",
         unemployed_pct = "Percent.Unemployed..2018.",
         senior_pct = "X65.",
         population_2018 = "Population.2018", ## extra variable in test
         non_elderly_pre_existing = "Nonelderly.Adults.Who.Have.A.Pre.Existing.Condition",
         homeless = "Number.Homeless..2019.",
         medical_expense = "Medicaid.Expenditures.as.a.Percent.of.Total.State.Expenditures.by.Fund",
         retail_recreation = 'Retail...recreation',
         transit = 'Transit.stations') %>%
  select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, 
         black_population_pct, white_population_pct, poverty_pct, unemployed_pct, senior_pct,
         population_2018, non_elderly_pre_existing, homeless, medical_expense, retail_recreation, transit)
# Assign correct data types
cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
df[cols.num] <- sapply(df[cols.num],as.numeric)
df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
head(df)
```

@@

### Model 3 Regression

Next, the regression model for model 3 is created based on the existing work from model 2. Two additional variables discussed in Model 3 will be introduced as maximalist approach. The first variable is the "mask_public_bool" (whether the state enforced public mask mandate or not) that represents the policy effects. The second variable introduced is "population_density" (the value is logged as discussed earlier) which is considered an important factor related to the spread of disease.

@@

```{r, warning=FALSE}

model3 <- lm(case_rate_100k ~ senior_per_100k 
              + poverty_per_100k 
              + unemployed_per_100k 
             + mask_public_bool
             + log(population_density), data = df)
coeftest(model3)

@@

### CLM Assumptions Validation for Model 3:

#### I.I.D Assumption

( Discussed in the common section)

@@

#### Normality of Errors

From the Normal Q-Q plot below, the points generally stay close to the normal line, with the tails exhibit certain degree of skew. There is no significant evidence of concerning normality in the errros.
```{r}
plot(model3, which=2)
```

#### Homoskedasticity

According to the plot of Homoskedasticity, the standardized residual falls into a reasonable range and the line looks smooth without concerning skews. There is no significant evidence that the model violates homoskedasticity assumptions.
```{r}

plot(model3, which=3)

#### Collinearity

Based on the Variance Inflation Factors check, all the variables have variance-inflation factors less than 4. As a result, there's no strong evidence of collinearity concern within the model. 
```{r}

vif(model3)>4
```

#### Cook's Distance

From the Cook's plot, there is no strong evidence of concerning outliers.

plot(model3, which=5)
```
